---
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
    toc: true
    toc_depth: 1
---

# Required Packages
In this project, we will use a number of amazing R packages:

  * `knitr`: to generate pretty tables of date using the `kable()` function.  
  * `tidyverse`: to manipulate data, generate plots (via `ggplot2`), and to use piping/chaining.
  * `rmarkdown`: to knit output files manually using the `render()` function.  
  * `jsonlite`: to pull data from various endpoints of the Covid 19 APIs.
  
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# Load necessary packages
library(knitr)
library(tidyverse)
library(rmarkdown)
library(jsonlite)
library(httr)
```


```{r}
#helper function to generate a vector of countries and their corresponding Slug name whose data are update daily.

countryList <- function(){
  covid <- GET("https://api.covid19api.com/summary")
  covidDF <- fromJSON(rawToChar(covid$content))
  countryList <- covidDF$Countries %>% select(Country, Slug) %>% as_tibble()
  return(countryList)
}
  
countryList <- countryList()
countryList
```

```{r}
#function to generate data of a particular country

countryData <- function(countrySlug){
  if (countrySlug %in% countryList$Slug){
      url <- paste0("https://api.covid19api.com/total/country/",countrySlug)
      covidData <- GET(url)
      covidDF <- fromJSON(rawToChar(covidData$content))
      dataTibble <- covidDF %>% as_tibble()
  } else {
    message <- paste("ERROR: Argument for country slug is not found.
                     Check the list of countries with its corresponding slug name
                     and use quotation marks.")
    stop(message)
  }
  return(dataTibble)
}

#now we can use the function to generate some data from five sample countries
covidInd <- countryData("indonesia")
covidBang <- countryData("bangladesh")
covidPhil <- countryData("philippines")
covidViet <- countryData("vietnam")
covidThai <- countryData("thailand")
```


# Data Manipulation
```{r}
#we can create a function that will manipulate our data to prepare for data summaries and visualization
manipulateData <- function(dataset){
  dataset$Date <- as.Date(dataset$Date)
  dataset <- dataset %>% select(-c(CountryCode, Province, City, CityCode, Lat, Lon)) %>%
                  separate(Date, c("Year", "Month", "Day"), sep="-", convert=T, 
                     remove=F) %>% mutate("NewCases"=diff(c(0, Confirmed)),
                                          "NewDeaths"=diff(c(0, Deaths)))
  dataset$Country <- as.factor(dataset$Country)
  return(dataset)
}

#manipulate each dataset using the function above to prepare for summarization and visualization
covidInd <- manipulateData(covidInd)
covidBang <- manipulateData(covidBang)
covidPhil <- manipulateData(covidPhil)
covidViet <- manipulateData(covidViet)
covidThai <- manipulateData(covidThai)


#histogram
ggplot(covidPhil, aes(NewDeaths)) + geom_histogram()

#barplot
dataBarPlot <- dataset %>% filter(Country=="Philippines", Year==2021, Month==9)
ggplot(dataBarPlot, aes(Day, NewDeaths)) + geom_col()




dataset <- bind_rows(covidBang, covidInd, covidPhil, covidThai, covidViet)


datasetPlot <- dataset %>% filter(Country %in% c("Indonesia", "Philippines", "Vietnam"))


ggplot(dataset, aes(Date,NewDeaths)) + geom_area(aes(fill=Country, alpha=0.5)) + geom_text(aes(label=NewDeaths))


#scatterplot
dataBarPlot <- dataset %>% filter(Country %in% c("Philippines", "Thailand"), Year==2021, Month==9)
ggplot(dataBarPlot, aes(Day, NewDeaths)) + geom_jitter(aes(color=Country))

#boxplot
dataBoxPlot <- dataset %>% filter(Country %in% c("Philippines", "Thailand"), Year==2021, Month==9)

ggplot(dataBoxPlot, aes(Country %in% c("Philippines", "Thailand"), NewDeaths)) + 
  geom_boxplot(aes(colour=Country))
```

```{r warning=F, message=F}

covidInd <- covidInd %>% filter(Year == 2021 & Month == 09)

ggplot(covidInd, aes(covidInd$Date, covidInd$NewCases)) + geom_col()

ggplot(covidInd, aes(covidInd$Date, covidInd$NewDeaths)) + geom_col()
```





```{r}
#get results of five countries to compare
compareCountries <- function(country1, country2, country3, country4, country5){
  covid <- GET("https://api.covid19api.com/summary")
  covidDF <- fromJSON(rawToChar(covid$content))
  data <- covidDF$Countries %>% as_tibble()
  #checks if first country supplied is in the data, misspelled or lack quotation marks
  if (country1 %in% data$Country){
    output1 <- data %>% filter(data$Country == country1) %>% select(-c(ID,Slug,Premium))
  }
  else{
    message <- paste("ERROR: Check spelling of first country or use quotation marks.")
    stop(message)
  }
  #checks if second country supplied is in the data, misspelled or lack quotation marks
  if (country2 %in% data$Country){
    output2 <- data %>% filter(data$Country == country2) %>% select(-c(ID,Slug, Premium))
  }
  else{
    message <- paste("ERROR: Check spelling of second country or use quotation marks.")
    stop(message)
  }
  #checks if third country supplied is in the data, misspelled or lack quotation marks
  if (country3 %in% data$Country){
  output3 <- data %>% filter(data$Country == country3) %>% select(-c(ID,Slug, Premium))
  }
  else{
    message <- paste("ERROR: Check spelling of third country or use quotation marks.")
    stop(message)
    }
  #checks if fourth country supplied is in the data, misspelled or lack quotation marks  
  if (country4 %in% data$Country){
  output4 <- data %>% filter(data$Country == country4) %>% select(-c(ID,Slug, Premium))
  }
  else{
    message <- paste("ERROR: Check spelling of fourth country or  use quotation marks.")
    stop(message)
  }
  #checks if fifth country supplied is in the data, misspelled or lack quotation marks  
  if (country5 %in% data$Country){
  output5 <- data %>% filter(data$Country == country5) %>% select(-c(ID, Slug, Premium))
  }
  else{
    message <- paste("ERROR: Check spelling of fifth country or use quotation marks.")
    stop(message)
  }
  output <- rbind(output1, output2, output3, output4, output5) %>% kable()
  return(output)
}

#check function if it works
compareCountries("Philippines","China", "Mexico", "United States of America", "Australia")

```


